---
title: "Circumference effect relative size selection"
output:
  word_document: default
  html_document: default
---

<!-- Modified from Expt 5 T90RelSeln.Rmd & Expt 6 T90DanishSeineRelSeln.Rmd -->

We have three treatments, Ninety-oneT0, Seventy-oneT90narrow and Seventy-oneT90wide.

Flatfish catches are subsampled in batches, so will need to be combined and use average SF.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.height=5)
```

```{r, echo=-(1:4),message=F,warning=F}
#remotes::install_local("T:/Russell_SSD/ACTIVE/CurrentWork/RPackages/SELECT",force=T)
#require(gridExtra)
#require(ggpubr)
#require(stringr) #For str_sub() to use -ve string posns 
require(SELECT)
require(mgcv)
require(tidyverse)
require(readxl)

```

## Input species size data and put into SELECT format
```{r, echo=F}
workdir="T:/Russell_SSD/ACTIVE/Collaboration/CTU/FRDC/Expt7T90&CircumEffectsOnFlathead/"
TF.df=read_excel(paste0(workdir,"Tigerflathead sizes for Russell.xlsx"),sheet="Data")
TF.df$Species="Flathead"
ESW.df=read_excel(paste0(workdir,"Eastern school whiting sizes for Russell.xlsx"),sheet=1)
ESW.df$Species="Whiting"
GUR.df=read_excel(paste0(workdir,"Round.snouted gurnard sizes for Russell.xlsx"),sheet=1)
GUR.df$Species="Gurnard"
All.df=rbind(TF.df,ESW.df,GUR.df)
All.df=rename(All.df,SF=ScalingFactor,Gear=Codend,lgth=Species.Size,
              TowID=Set.Number,nSamp=TotalNumberofFishMeasured,
              nTot=NumberofFishinCatch,n=NumberMeasuredatLength)
#All.df %>% group_by(Gear) %>% summarize(nTows=length(unique(TowID)))
All.df %>% group_by(Day,TowID) %>% summarize(Gear=unique(Gear)) |> print(n=40)
#Only days 4(4), 14(4), 9(2), 6(3) & 13(3) have hauls with both T90N and T90W 

#Shorten Gear names
All.df$Gear=c("T0","T90N","T90W")[match(All.df$Gear,c("90-mm","Narrow-T90","Wide-T90"))]
All.df$q=1/All.df$SF
All.df=subset(All.df,select=-c(WeightedNumber,Fate,nTot,SF,nSamp)) %>% data.frame()
#All.df$Row=1:nrow(All.df)

DF=SELECT.FORMAT(All.df,by=c("Species","Day","TowID","lgth"),
                          gear="Gear",freq="n",q.name="q",paired=F)
```

```{r}
species=c("Flathead","Whiting","Gurnard") 
Spec=1 #Species selection
Gears=c("T0","T90N","T90W")
#SpecDf=DF %>% filter(Species==species[Spec], Gear!="T0")
SpecDf=DF %>% filter(Species==species[Spec], Gear!="T90N")

#SpecDf %>% group_by(Gear) %>% summarize(nTows=length(unique(TowID)))

var.names=c("lgth","nT90N","nT90W")
q.names=c("qT90N","qT90W")
var.names=c("lgth","nT0","nT90W")
q.names=c("qT0","qT90W")

Tots=Raw2Tots(SpecDf,var.names,q.names); 
#Calculate proportion in T90W include on plots
#Tots=Tots |> mutate(n=nT90N+nT90W,y=nT90W/n) |> filter(n>0) #
Tots=Tots |> mutate(n=nT0+nT90W,y=nT90W/n) |> filter(n>0) #

LOGLIKS(SpecDf,var.names,q.names)

lgth.seq=seq(15,60,0.5) #Length sequence for predictions
SplineFit=SplineSELECT(SpecDf,var.names,q.names,quasi=T)
Fitted=predict(SplineFit,newdata=data.frame(lgth=lgth.seq),type="response")
plot(lgth.seq,Fitted,type="l",xlab="Length (cm)",ylab="Catch share",
     main=paste(species[Spec],"T90N and T90W"),ylim=c(0,1))
points(Tots$lgth,Tots$y,pch=16)
```


## Bootstrap

```{r Bootstrap, echo=F}
#Define function that returns predictions (copy and paste above code)
Predn=function(data,var.names) {
  SplineFit=SplineSELECT(data,var.names,q.names,quasi=T)
  predict(SplineFit,newdata=data.frame(lgth=lgth.seq),type="response") 
}

nboot=400 #Number of bootstrap replicates
BootPredns=bootSELECT(SpecDf,var.names,Predn,haul="TowID",nsim=nboot,
                      gear="Gear",paired=F)

BootPlot(BootPredns,lgth.seq,predn=Fitted,Data=Tots)
```

## Permutation test

### Define useful permutation statistics

```{r Permutation, echo=F, warning=F}
#Define function that returns an appropriate test statistic(s)
permLRT=function(data,var.names,q.names) { #No good with quasi=T)
  SplineFit=SplineSELECT(data,var.names,q.names=NULL,quasi=T)
  LogLiks=LOGLIKS(data,var.names,fixed=c(0.5,0.5))
  LRT=2*(logLik(SplineFit)-LogLiks$null)
  LRT0.5=2*(logLik(SplineFit)-LogLiks$fixed) 
  c(LRT,LRT0.5) }
permRsq=function(data,var.names,q.names) {
  SplineFit=SplineSELECT(data,var.names,q.names=NULL,quasi=T)
  summary(SplineFit)$r.sq }
permCheck=function(data,var.names,q.names) {
  Wk=data %>% group_by(Day,TowID) %>% summarize(Gear=unique(Gear)) 
  Wk$Gear }
```

### Check the permutation functionality
```{r}
PermDf=Randomize(data=SpecDf, var.names[2:3], haul="TowID",
                     paired=F, gear="Gear", q.names=q.names) #, block="Day")
SpecDf %>% group_by(Day,TowID) %>% summarize(Gear=unique(Gear)) |> print(n=40)
PermDf %>% group_by(Day,TowID) %>% 
  summarize(Permuted=unique(Permuted),Gear=unique(permuted.Gear)) |> print(n=40)
```



permFunc=permRsq
#permFunc=permCheck

SpecDf %>% group_by(Day,TowID) %>% summarize(Gear=unique(Gear)) |> print(n=40)


obsStat=permFunc(SpecDf,var.names,q.names)
  
nperm=400
permStats=permSELECT(SpecDf,var.names,permFunc,haul="TowID",nsim=nperm,
                     paired=F,gear="Gear",q.names=q.names,block="Day")

permPval(obsStat,permStats)

#Arguments for Randomize, for testing


```

