---
title: 'Next Generation SELECT  \newline Examples of Use'
author: "Russell Millar"
output:
  word_document: default
  pdf_document: default
  html_document:
    fig_caption: yes
    number_sections: yes
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.height=3)
```

```{r}
remotes::install_github("rbmillar/SELECT")
require(SELECT)
require(tidyverse)
```

# Analysis of Clark's covered codend haddock data

**Logistic and Richards curves are fitted. We see that the logistic is preferred**

`Clark` **is a dataframe included within the SELECT package.**

```{r}
#Use dtype="cc" to specify that design type is covered codend
data(Clark)
head(Clark)
cc.names=c("lenclass","cover","codend")
#Fit logistic selection curve
fitL=SELECT(Clark,cc.names,dtype="cc")
ModelCheck(fitL,minE=1,las=1)
Estimates(fitL) 
PlotCurves(fitL,plotlens=seq(19,56,0.1))

#Fit Richards selection curve
fitR=SELECT(Clark,cc.names,dtype="cc",stype="richards")
ModelCheck(fitR,minE=1,las=1)
Estimates(fitR);
PlotCurves(fitR,plotlens=seq(19,56,0.1))

#Likelihood ratio test for H0 that logistic mode is adequate
LRTstat=2*(deviance(fitL)-deviance(fitR))
cat("\n p-value for H0:logistic is",1-pchisq(LRTstat,1))
```

**The `ModelCheck` shows no evidence of overdispersion so we can use twice the
difference in deviance as a likelihood ratio test statistic. 
There in no significant evidence against the logistic curve.**

# Analysis of Pope's alternate haul haddock data

```{r}
#Use dtype="ph" to specify that design type is paired haul
data(Pope)
head(Pope)
ph.names=c("Lengths","nfine","nwide")
#FIt logistic seln cure
fitL=SELECT(Pope,ph.names,dtype="ph")
ModelCheck(fitL,minE=1,las=1)
Estimates(fitL) 
PlotCurves(fitL)
```

# Analysis of Pope's alternate haul haddock data with fixed split (of 0.5)

**For fixed split, the user must specify start values for parameters of the selection curve.**

**The fixed split is specified using relative power, in this case, `rel.power=c(1,1)` since the experimental and control have the same fishing power when the split is 0.5.**

**Use `dtype=re` for `relative` (power) fits.**

```{r}
#Fixed split, psplit=0.5
fitR2=SELECT(Pope,ph.names,dtype="re",x0=c(-30,1),rel.power=c(1,1))
ModelCheck(fitR2,minE=1,las=1)
Estimates(fitR2) 
PlotCurves(fitR2,las=1)
```

# Analysis of Holt's gillnet data

**Several alternative curve types are fitted. The bi-lognormal looks preferable, evidenced by a decrease in deviance of about 160.**
**Even taking in to account the over-dispersion correction factor of about 9, this is still highly significant.**
**The starting parameters for the gear with smallest mesh size need to be specified.**

```{r}
data(Holt)
Meshs=Holt$Meshsize 
Counts=Holt$Counts
names(Counts)=c("lgth",paste0("M",Meshs))
re.names=names(Counts)
#Equal fishing power of the eight meshsizes
pwr=rep(1,8)
#Or use this for fishing power proportional to meshsize
#pwr=Meshsize

par(mfrow=c(3,2),mar=c(4.1,4.1,1,1))
fit=SELECT(Counts,re.names,dtype="re",stype="norm.loc",Meshsize=Meshs,x0=c(60,4),rel.power=pwr)
ModelCheck(fit,minE=1); Estimates(fit); PlotCurves(fit,plotlens=seq(40,90,0.1))

fit=SELECT(Counts,re.names,dtype="re",stype="norm.sca",Meshsize=Meshs,x0=c(60,4),rel.power=pwr)
ModelCheck(fit,minE=1); Estimates(fit); PlotCurves(fit,plotlens=seq(40,90,0.1))

#Needs wide seln curve for x0
fit=SELECT(Counts,re.names,dtype="re",stype="gamma",Meshsize=Meshs,x0=c(10,5),rel.power=pwr) 
ModelCheck(fit,minE=1); Estimates(fit); PlotCurves(fit,plotlens=seq(40,90,0.1))

fit=SELECT(Counts,re.names,dtype="re",stype="logistic",Meshsize=Meshs,x0=c(-60,1),rel.power=pwr) 
ModelCheck(fit,minE=1); Estimates(fit); PlotCurves(fit,plotlens=seq(40,90,0.1))

fit=SELECT(Counts,re.names,dtype="re",stype="richards",Meshsize=Meshs,x0=c(-60,1,0),rel.power=pwr) 
ModelCheck(fit,minE=1); Estimates(fit); PlotCurves(fit,plotlens=seq(40,90,0.1))

fit=SELECT(Counts,re.names,dtype="re",stype="lognorm",Meshsize=Meshs,x0=c(log(60),0.1),rel.power=pwr)
ModelCheck(fit,minE=1); Estimates(fit); PlotCurves(fit,plotlens=seq(40,90,0.1))

fit=SELECT(Counts,re.names,dtype="re",stype="binorm.sca",Meshsize=Meshs,x0=c(55,4,65,4,2),rel.power=pwr)
ModelCheck(fit,minE=1); Estimates(fit); PlotCurves(fit,plotlens=seq(40,90,0.1))

fit=SELECT(Counts,re.names,dtype="re",stype="bilognorm",Meshsize=Meshs,x0=c(4,0.2,4.2,0.1,2),rel.power=pwr)
ModelCheck(fit,minE=1); Estimates(fit); PlotCurves(fit,plotlens=seq(40,90,0.1))
```
