---
title: "SELECT package demonstration   \n   Gillnet"
output: word_document
editor_options: 
  markdown: 
    wrap: 80
---

## Summary

**This case study demonstrates:**

-  Fitting six different shapes of selection curves to gillnet data
-  Use of `dtype=dc` for direct comparison experiments whereby relative fishing
power is NOT being estimated. (The default is equal power, but it can be specified by the user.)
-  Model selection using AIC
-  Geometric similarity of selectivity curves

### Data source

The data are for trout caught by gillnet in Gull Island Refuge, Lake Superior
(Hansen et al., 1997. Gillnet selectivity for lake trout (Salvelinus namaycush)
in Lake Superior. Canadian Journal of Fisheries and Aquatic Sciences, 54:
2483-2490.

```{r, echo=F}
knitr::opts_chunk$set(fig.height = 3.5)
```

### Load required packages

```{r, warning=F, message=F, echo=-1}
#devtools::install_github("rbmillar/SELECT")
require(SELECT)
require(tidyverse)
```

### Load data

```{r}
data(Trout)
Trout
```

Fit five different parametric selectivity curves to the data.

1.  Normal with common standard deviation (spread)
2.  Normal with standard deviation scaling with mesh size
3.  Log normal
4.  Binormal mixture
5.  Bilognormal mixture

Note that curves 2-5 all observe geometric similarity whereby the curves scale
in proportion to mesh size.

### Fit the curves

```{r}
Meshs=Trout$Meshsize
Counts=Trout$Counts 
Vars=names(Counts)

SEL1=SELECT(Counts,Vars,dtype="dc",
stype="norm.loc",Meshsize=Meshs) 
AIC1=AIC(SEL1,type="Poisson")

SEL2=SELECT(Counts,Vars,dtype="dc",
stype="norm.sca",Meshsize=Meshs) 
AIC2=AIC(SEL2,type="Poisson")

SEL3=SELECT(Counts,Vars,dtype="dc", stype="gamma",Meshsize=Meshs)
AIC3=AIC(SEL3,type="Poisson")

SEL4=SELECT(Counts,Vars,dtype="dc", stype="lognorm",Meshsize=Meshs)
AIC4=AIC(SEL4,type="Poisson")

SEL5=SELECT(Counts,Vars,dtype="dc",
stype="binorm.sca",Meshsize=Meshs)
AIC5=AIC(SEL5,type="Poisson")

SEL6=SELECT(Counts,Vars,dtype="dc",
stype="bilognorm",Meshsize=Meshs)
AIC6=AIC(SEL6,type="Poisson")

cat("Poisson AICs are",AIC1,AIC2,AIC3,AIC4,AIC5,AIC6)
```

The above SELECT fits show that the bimodal selectivity curves are preferred, and
the binormal (`SEL5`) has the lowest Poisson AIC of 412.9.

### Check goodness of fit of preferred model
```{r}
ModelCheck(SEL5)
cat("Goodness of fit  p-value is",1-pchisq(80.6,67))           
```

The above output from `ModelCheck`` enabled a goodness-of-fit test using the Pearson chi-square statistic (or deviance), and this shows that the Poisson assumption is
reasonable. The plot of deviances residuals does not dispay any concerning patterns.

### Plot the master selection curve

Due to geometric similarity, for each mesh size the selection curve scales (as a function of length) in proportion to mesh size, so the master curve describes selectivity for all mesh sizes.

```{r, fig.width=7, fig.height=4.5}
par(oma=c(0,0,0,0),mar=c(4,4,1,4)) 
plot(SEL5,standardize=T) #Use standardize=T to set maximum retention to 1
#Add plot legend
legend(600,0.3,legend=paste(Meshs,"mm"),lty=1:5,col=1:5,cex=0.8)
```


