---
title: "SELECT package demonstration   \n   Paired-haul relative selectivity "
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r, echo=F}
knitr::opts_chunk$set(fig.height = 3.5)
```

### Data source

The data are for school prawn relative selectivity in penaeid trawls with differing mesh types of the side panel, 43 mm diamond, or 32 or 35 mm square (Broadhurst et al., 2018. T45 side panels improve penaeid-trawl selection. Fisheries Research, 204: 8-15). 

### Load required packages

```{r,warning=FALSE,echo=F}
devtools::install( paste0("T:/Russell_SSD/ACTIVE/CurrentWork/RPackages/SELECT") )
require(tidyverse)
require(mgcv)
require(SELECT)
require(readxl) #This package is installed with tidyverse
```

## Read in data


NOTE: The last paired haul (D6H4) of Conv-T34 has NA for paenaeid counts in 
the conventional gear and so must be removed 

```{r}
Conv.df=read_excel("Exp 18 LenFreqs.xlsx", sheet = "Conventional trawl", na="NA")
T32.df=read_excel("Exp 18 LenFreqs.xlsx", sheet = "32 square trawl", na="NA")
T34.df=read_excel("Exp 18 LenFreqs.xlsx", sheet = "34 square trawl", na="NA")

Conv.df = Conv.df |> rename(Haul=Day, nConv=No.school, sfConv=Sf.school)
T32.df = T32.df |> rename(Haul=Day, n32=No.school, sf32=Sf.school)
T34.df = T34.df |> rename(Haul=Day, n34=No.school, sf34=Sf.school)
#Conv.df = Conv.df |> rename(Haul=Day, nConv=No.king, sfConv=Sf.king)
#T32.df = T32.df |> rename(Haul=Day, n32=No.king, sf32=Sf.king)
#T34.df = T34.df |> rename(Haul=Day, n34=No.king, sf34=Sf.king)
#The D6H4 haul pair of Conv vs T34 has missing counts for the conventional gear. 
#Remove haul pair D6H4 for these two gears.
Conv.df = Conv.df |> filter(Haul != "D6H4")
T34.df = T34.df |> filter(Haul != "D6H4")
```


## Create SELECT format dataframes for paired tows with conventional gear

Note the conversion from sub-sampling scaling factors to sampling fractions
and removal of CLs that are outside of the range of measured data.

```{r}
#Conventional and 32mm side gears
Conv32.df = inner_join(Conv.df,T32.df,by=c("Haul","CL"))
#Convert scaling factors to sampling fractions
Conv32.df = Conv32.df |>
   transform(qConv=1/sfConv,q32=1/sf32) |> filter(CL>5 & CL<28)
#Conventional and 35mm side gears
Conv34.df = inner_join(Conv.df,T34.df,by=c("Haul","CL"))
#Convert scaling factors to sampling fractions
Conv34.df = Conv34.df |>
   transform(qConv=1/sfConv,q34=1/sf34) |> filter(CL>5 & CL<28)
```

## Spline relative selection curve for 32 mm T45
```{r}
var.names=c("CL","nConv","n32")
q.names=c("qConv","q32")
T32.spline=SplineSELECT(Conv32.df,var.names,q.names,bs="cr",quasi=T,sumHauls=T,
                        k=5,q.ODadjust = T)
summary(T32.spline)
```

#Calculate over-dispersion
```{r}
yhat=fitted(T32.spline)
Catch=Raw2Tots(Conv32.df,var.names,q.names,sumHauls=F)
N=Catch$nConv+Catch$n32
O=Catch[,c("nConv","n32")]; E=N*cbind(yhat,1-yhat)
calcOD(O,E,5)
```

```{r Define bootstrap funcs}
#Define the bootstrap prediction function
CLseq=seq(5,25,0.5) #Carapace lengths to use for predn
Predn=function(data,var.names) {
  SplineFit=SplineSELECT(data,var.names,q.names,bs="cr",quasi=T,sumHauls=F,k=5,
                         q.ODadjust = T)
  predict(SplineFit,newdata=data.frame(CL=CLseq),type="response") }
#Check it works
Tot32.df=Raw2Tots(Conv32.df,var.names,q.names) |> 
     transform(lgth=CL, y=n32/(nConv+n32))
plot(y~CL,data=Tot32.df,ylim=c(0,1),xlab="Carapace length (mm)",
     ylab="T32 catch share",main="32 mm T45")
points(CLseq,yhat,type="l")
abline(h=0.5,lty=3)
```

## Do the bootstrap

```{r, warning=F, message=F}
nboot=1000 #Increase to at least 1000 in practice
Boot32=bootSELECT(Conv32.df,var.names,Predn,haul="Haul",nsim=nboot,paired=T)
```

```{r}
BootPlot(Boot32,CLseq,yhat,Data=Tot32.df)
```

## Do permutation test
```{r}
#Define the spline r-squared function
permRsq=function(data,var.names) {
  SplineFit=SplineSELECT(data,var.names,q.names,bs="ts",quasi=T,sumHauls=F,k=5,
                         q.ODadjust = T)  
  summary(SplineFit)$r.sq }
#Check it works
T32.rsq=permRsq(Conv32.df,var.names)
```

```{r, warning=F, message=F}
nperm=1000 #Increase to at least 1000 in practice
Perm32=permSELECT(Conv32.df,var.names,permRsq,haul="Haul",nsim=nperm,paired=T)

mean(Perm32>T32.rsq) #Proportion of permuted r-squared values greater than the observed
```


## Redo for 35 mm T45


## Spline relative selection curve for 34 mm T45
```{r}
var.names=c("CL","nConv","n34")
q.names=c("qConv","q34")
T34.spline=SplineSELECT(Conv34.df,var.names,q.names,bs="cr",quasi=F,sumHauls=F,
                        k=5,q.ODadjust = T)
```




```{r Define bootstrap funcs}
#Define the boot function
CLseq=seq(5,28,0.5) #Carapace lengths to use for predn
Predn=function(data,var.names) {
  SplineFit=SplineSELECT(data,var.names,q.names,bs="cr",quasi=T,sumHauls=F,k=11,
                         q.ODadjust = T)
  predict(SplineFit,newdata=data.frame(CL=CLseq),type="response") }
#Check it works
yhat=Predn(Conv34.df,var.names)
Tot34.df=Raw2Tots(Conv34.df,var.names,q.names) |> 
     transform(lgth=CL, y=n34/(nConv+n34))
plot(y~CL,data=Tot34.df,ylim=c(0,1),xlab="Carapace length (mm)",
     ylab="T34 catch share",main="34 mm T45")
points(CLseq,yhat,type="l")
abline(h=0.5,lty=3)
```

## Do the bootstrap

```{r, warning=F, message=F}
nsim=1000 #Increase to at least 1000 in practice
Boot34=bootSELECT(Conv34.df,var.names,Predn,haul="Haul",nsim=nsim,paired=T)
```

```{r}
BootPlot(Boot34,CLseq,yhat,Data=Tot34.df)
```
 
## Do permutation test
```{r}
T34.rsq=permRsq(Conv34.df,var.names)
```

```{r, warning=F, message=F}
nperm=1000 #Increase to at least 1000 in practice
Perm34=permSELECT(Conv34.df,var.names,permRsq,haul="Haul",nsim=nperm,paired=T)

mean(Perm34>T34.rsq) #Proportion of permuted r-squared values greater than the observed
```
