---
title: "SELECT package demonstration   \n   Paired-haul relative selectivity "
output:
  word_document: default
  html_document: default
  pdf_document: default
---

## Summary

**This case study demonstrates:**

-   Reading data from separate gear-specific files and merging them into a
    single data frame.
-   Handling sub-sampled data.
-   Fitting splines to relative selectivity data using the `SplineSELECT` function.
-   Using the bootstrap function `bootSELECT` to estimate the uncertainty in the
    catch share curve.
-   Using the permutation function `permSELECT` to test for a difference in the
    size selectivity of the two gears.

```{r, echo=F}
knitr::opts_chunk$set(fig.height = 3.5)
```

### Data source

The data are for school prawn relative selectivity in penaeid trawls with differing mesh types of the side panel, 43 mm diamond, or 35 mm square (Broadhurst et al., 2018. T45 side panels improve penaeid-trawl selection. Fisheries Research, 204: 8-15). 

### Load required packages

```{r,warning=FALSE,message=F,echo=-1}
#devtools::install( paste0("T:/Russell_SSD/ACTIVE/CurrentWork/RPackages/SELECT") )
require(tidyverse)
require(mgcv)
require(SELECT)
require(readxl) #This package is installed with tidyverse
```

### Read in data
  
NOTE: Paired haul D6H4 with conventional/T35 pairing has NAs for the
penaeid counts in the conventional gear, and so must be removed 

```{r}
Conv.df=read_excel("Exp 18 LenFreqs.xlsx", sheet = "Conventional trawl", na="NA")
T35.df=read_excel("Exp 18 LenFreqs.xlsx", sheet = "35 square trawl", na="NA")

Conv.df = Conv.df |> rename(Haul=Day, nConv=No.school, sfConv=Sf.school)
T35.df = T35.df |> rename(Haul=Day, n35=No.school, sf35=Sf.school)
Conv.df = Conv.df |> filter(Haul != "D6H4")
T35.df = T35.df |> filter(Haul != "D6H4")
```

### Merge the separate dataframes to create a dataframe for the paired tows 

Note the conversion from sub-sampling scaling factors to sampling fractions
and removal of CLs that are outside of the range of measured data.

```{r}
#Merge conventional and 35mm side gears
Conv35.df = inner_join(Conv.df,T35.df,by=c("Haul","CL"))
#Convert scaling factors to sampling fractions
Conv35.df = Conv35.df |>
   transform(qConv=1/sfConv,q35=1/sf35) |> filter(CL>5 & CL<26)
```

### Define variable names
```{r}
var.names=c("CL","nConv","n35")
q.names=c("qConv","q35")
```

```{r, echo=F, eval=F}
#Calculate over-dispersion
T35.spline=SplineSELECT(Conv35.df,var.names,q.names,bs="cr",quasi=T,sumHauls=T,
                        k=5,q.ODadjust = F)
yhat=fitted(T35.spline)
Catch=Raw2Tots(Conv32.df,var.names,q.names,sumHauls=F,q.ODadjust = F)
N=Catch$nConv+Catch$n32
O=Catch[,c("nConv","n32")]; E=N*cbind(1-yhat,yhat)
calcOD(O,E,5.473,minE=3)
#cf sum(resid(T35.spline, type="pearson")^2)
```

###  Define a prediction function to be used with the bootstrap
```{r}
#Define the bootstrap prediction function
CLseq=seq(5,25,0.5) #Carapace lengths to use for predn
Predn=function(data,var.names) {
  SplineFit=SplineSELECT(data,var.names,q.names,bs="cr",quasi=T,sumHauls=F,k=5,
                         q.ODadjust = T)
  predict(SplineFit,newdata=data.frame(CL=CLseq),type="response") }
#Check it works
Tot35.df=Raw2Tots(Conv35.df,var.names,q.names) |> 
     transform(lgth=CL, y=n35/(nConv+n35))
plot(y~CL,data=Tot35.df,ylim=c(0,1),xlab="Carapace length (mm)",
     ylab="T35 catch share",main="35 mm T45")
predn=Predn(Conv35.df,var.names)
points(CLseq,predn,type="l")
abline(h=0.5,lty=3)
```

### Do the bootstrap

```{r, warning=F, message=F}
nboot=100 #Increase to at least 1000 in practice
Boot35=bootSELECT(Conv35.df,var.names,Predn,haul="Haul",nsim=nboot,paired=T)
```

```{r}
BootPlot(Boot35,CLseq,predn,Data=Tot35.df)
```

###  Define an r-squared function to be used with the permutation test

```{r}
#Define the spline r-squared function
permRsq=function(data,var.names) {
  SplineFit=SplineSELECT(data,var.names,q.names,bs="ts",quasi=T,sumHauls=F,k=5,
                         q.ODadjust = T)  
  summary(SplineFit)$r.sq }
#Check it works
T35.rsq=permRsq(Conv35.df,var.names)
```

```{r, warning=F, message=F}
nperm=100 #Increase to at least 1000 in practice
Perm35=permSELECT(Conv35.df,var.names,permRsq,haul="Haul",nsim=nperm,paired=T)
#Proportion of permuted r-squared values greater than the observed
cat("The p-value for a CL effect is",mean(Perm35>T35.rsq),"\n")
```
