---
title: "SELECT package demonstration   \n   Relative selectivity "
output:
  pdf_document: default
  html_document: default
  word_document: default
---

\small

NOTE: The last paired haul (D6H4) of Conv-T34 has NA for paenaeid counts in Conv gear 
- must be removed if fitting to totals.

### Data source

The data are for school prawn relative selectivity in penaeid trawls with differing mesh types of the side panel, 43 mm diamond, or 32 or 35 mm square (Broadhurst et al., 2018. T45 side panels improve penaeid-trawl selection. Fisheries Research, 204: 8-15). 

```{r,warning=FALSE,echo=F}
#Load required packages
require(tidyverse)
require(mgcv)
require(SELECT)
require(XLConnect)
#require(data.table)
#library(dplyr)
#require(lme4)
#require(gamm4)
#source("R:/Russell_HD/CTU/BootstrapDf.R")
```

## Read in data
```{r}
wb <- loadWorkbook("Exp 18 LenFreqs.xlsx", create = FALSE)
summary(wb)
setMissingValue(wb,value="NA")
Conv.df = readWorksheet(wb,sheet = "Conventional trawl")
T32.df = readWorksheet(wb,sheet = "32 square trawl")
T34.df = readWorksheet(wb,sheet = "34 square trawl")
Conv.df = Conv.df |> rename(Haul=Day, nConv.S=No.school, sfConv.S=Sf.school,
                            nConv.K=No.king, sfConv.K=Sf.king) 
T32.df = T32.df |> rename(Haul=Day, n32.S=No.school, sf32.S=Sf.school,
                            n32.K=No.king, sf32.K=Sf.king)  
T34.df = T34.df |> rename(Haul=Day, n34.S=No.school, sf34.S=Sf.school,
                            n34.K=No.king, sf34.K=Sf.king)
```



```{r}
#Create dataframes for paired tows having conventional gear
#Conventional and 32mm side gears
Conv32.df = inner_join(Conv.df,T32.df,by=c("Haul","CL"))
#Convert scaling factors to sampling fractions
Conv32.df = Conv32.df |>
   transform(qConv.S=1/sfConv.S,qConv.K=1/sfConv.K,
             q32.S=1/sf32.S,q32.K=1/sf32.K) |> filter(CL>5 & CL<26)
#Conventional and 35mm side gears
Conv34.df = inner_join(Conv.df,T34.df,by=c("Haul","CL"))
Conv34.df = Conv34.df |> filter(Haul != "D6H4")
#Convert scaling factors to sampling fractions
Conv34.df = Conv34.df |>
   transform(qConv.S=1/sfConv.S,qConv.K=1/sfConv.K,
             q34.S=1/sf34.S,q34.K=1/sf34.K) |> filter(CL>5 & CL<26)
```




```{r Define funcs}
###Set up the boot function and the r-squared function
CLseq=seq(5,25,1) #Sequence to use for predn
Predn=function(data,var.names) {
  SplineFit=SplineSELECT(data,var.names,q.names,bs="cr",quasi=T,Tots=F,k=5,q.ODadjust=F)
  predict(SplineFit,newdata=data.frame(CL=CLseq),type="response") }
permRsq=function(data,var.names,q.names,k=4) {
  SplineFit=SplineSELECT(data,var.names,q.names,k=k)
  summary(SplineFit)$r.sq }
```


# Conventional vs 32 mm T45

```{r}
#Schoolies
var.names=c("CL","nConv.S","n32.S")
q.names=c("qConv.S","q32.S")
TotDf=Raw2Tots(Conv32.df,var.names,q.names) |> 
    transform(lgth=CL,y=n32.S/(nConv.S+n32.S))
plot(y~lgth,data=TotDf,ylim=c(0,1),xlab="Carapace length (mm)")
yhat=Predn(Conv32.df,var.names); yhat
plot(CLseq,yhat)
```

##Bootstrap

```{r, warning=F, message=F}
Conv32.Out=bootSELECT(Conv32.df,var.names,Predn,haul="Haul",nsim=1000,
                    paired=T)
```

```{r}
BootPlot(Conv32.Out,CLseq,yhat,Data=TotDf)
```

# Conventional vs 34 mm T45 (To be completed)

```{r}
#Schoolies
var.names=c("CL","nConv.S","n34.S")
q.names=c("qConv.S","q34.S")
TotDf=Raw2Tots(Conv34.df,var.names,q.names) |> 
    transform(lgth=CL,y=n34.S/(nConv.S+n34.S))
plot(y~lgth,data=TotDf,ylim=c(0,1),xlab="Carapace length (mm)")
yhat=Predn(Conv34.df,var.names); yhat
plot(CLseq,yhat)
```
##Bootstrap

```{r, warning=F, message=F}
Conv34.Out=bootSELECT(Conv34.df,var.names,Predn,haul="Haul",nsim=1000,
                    paired=T)
```
```{r}
BootPlot(Conv34.Out,CLseq,yhat,Data=TotDf)
```
