---
title: "SELECT package demonstration   \n   Alternate haul"
output: word_document
---

## Summary

This is a simple case study that revisits the alternate haul data analysed
in Wileman et al. (1996, Manual of Methods of Measuring the Selectivity of Towed
Fishing Gears, ICES Coop Research Report, No 215).

**This case study demonstrates:**

-   The use of data that is provided within the `SELECT` package.
-   Fixing the relative fishing power using the direct comparison 
experimental design type `dtype=dc`.

### Data source

The data are originally from Pope et al. (1975, Manual of methods for fish stock
assessment Part 111 Selectivity of fishing gear.FA0 Fish. Tech. Pap. (41) Rev. 1.)
These are alternative haul catch data of haddock in an 87 mm diamond experimental
codend alternating with a 35 mm control gear.

The data are only given as totals summed over hauls, so no exploration of
between-haul variability is possible. The data are included with the `SELECT`
package.

```{r, echo=F}
knitr::opts_chunk$set(fig.height = 3.5)
```

```{r, warning=F, message=F}
#devtools::install_github("rbmillar/SELECT")
require(SELECT)
require(tidyverse)
```

### Load and inspect the data
```{r}
data(Pope)
Pope
```

### Define variable names
```{r}
v.names=c("Lengths","nfine","nwide")
```

### Produce a plot of catch-share proportions
```{r}
Pope=transform(Pope,n=nfine+nwide,y=nwide/(nfine+nwide)) 
plot(y~Lengths,data=Pope)
abline(h=0.5,lty=3)
```

The experimental condend has catch share of 50% or more for all of the larger 
lengths, indicating that it has higher fishing power than the control.

### Fit logistic selection curve
```{r}
Logist.fit=SELECT(Pope,var.names=v.names,dtype="ph")
plot(Logist.fit)
ModelCheck(Logist.fit)
Estimates(Logist.fit)
```
The logistic fit looks good. For completeness we'll do the Richards fit also.

### Fit Richards curve
```{r}
Rich.fit=SELECT(Pope,var.names=v.names,dtype="ph",stype="richards")
plot(Rich.fit)
ModelCheck(Rich.fit)
Estimates(Rich.fit)
```

There is very little difference between the logistic and Richards fits.

The logistic is preferred, since even without being able to correct for 
over-dispersion, AIC prefers the logistic. A likelihood-ratio test also
would not reject goodness of fit of the logistic.


## Fixing the split parameter

Fixing the value of the split parameter requires specifying the design type 
as direct comparison, `dc`.
Also, meshsizes must be given.

The default is equal fishing power, i.e., 50/50 split. 
The relative fishing power is specified through a vector of 
length two that is provided as the `rel.power` argument. 

```{r, warning=F}
#50/50 split
EqualPower.fit=SELECT(Pope,var.names=v.names,dtype="dc",
                      Meshsize=c(3.5,8.7),rel.power=c(0.5,0.5))
plot(EqualPower.fit)
ModelCheck(EqualPower.fit)
Estimates(EqualPower.fit)
```

The equal fishing power model is clearly not a good fit.



